---
title: "analysis_challenger"
output: html_document
date: "2025-01-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r libraries}
library(tidyverse)
library(ggplot2)
```

## Loading data

```{r loading data}
data <- read.csv("../data/shuttle.csv", header = TRUE)
```

### Adding frequency of O-rings failure
```{r failure frequency}
data$Frequency <- data$Malfunction / data$Count
datarm <- data %>%
  filter(Malfunction > 0)
```

### Scatter plot of Temperature vs Frequency of failure
```{r scatter plot}
# Add a new column to identify the dataset
data$Source <- "All Data"
datarm$Source <- "Filtered Data"

# Combine the two datasets into one
combined_data <- bind_rows(data, datarm)

# Create the plot with facet_wrap to display both datasets side by side
ggplot(combined_data, aes(x = Temperature, y = Frequency)) +
  geom_point() +
  labs(title = "Temperature vs O-ring Failure Frequency",
       x = "Temperature (°F)",
       y = "Failure Frequency") +
  theme_minimal() +
  facet_wrap(~ Source, scales = "free_y")  # Display two panels with separate y-axis for each dataset

```



From these graph we can see that for the non filtered data there seem to be a trend regarding temperature which is imperceptible with the filtered data.

## Logistic regression only with temperature
```{r logistic regression temperature alone}
logmodel <- glm(Frequency ~ Temperature, 
                data = data, 
                family = binomial(link = "logit"))
logmodelrm <- glm(Frequency ~ Temperature, 
                data = datarm, 
                family = binomial(link = "logit"))
summary(logmodel)
summary(logmodelrm)
```

In either of these data set (filtered vs unfiltered) temperature doesn't show significance with a P value = 0.5 for the unfiltered data and 0.86 for the filtered data, at first this could suggest that removing data doesn't have an impact in the significance of the temperature. However in this analysis the pressure which is part of this dataset has not been taken into account

## Logistic regression with temperature and pressure
```{r log reg temp + press}
model_2 <- glm(cbind(Malfunction, Count - Malfunction) ~ Temperature + Pressure, 
               data = data, 
               family = binomial(link = "logit"))

model_2rm <- glm(cbind(Malfunction, Count - Malfunction) ~ Temperature + Pressure, 
               data = datarm, 
               family = binomial(link = "logit"))

summary(model_2)
summary(model_2rm)

```

When we apply the logistic regression considering temperature and pressure, the analysis reveals that temperature has a significant impact on the probability of O-ring failure with the full dataset with a p value = 0.029. The negative coefficient suggest that as temperature increases, the probability of failure decreases.

However when the same model is applied to the filtered dataset, temperature doesn't show any significance. This shift suggest that removing datas reduced the model's ability to capture meaningfull information leading to a loss of critical information.

So far we've tried to include temperature and pressure and since the original dataset doesn't include any other variable, the only thing left we could explore is the interaction between these two variables.

## logistic regression of temperature * pressure
```{r temperature*pressure}
model_3 <- glm(cbind(Malfunction, Count - Malfunction) ~ Temperature * Pressure, 
               data = data, 
               family = binomial(link = "logit"))

model_3rm <- glm(cbind(Malfunction, Count - Malfunction) ~ Temperature * Pressure, 
               data = datarm, 
               family = binomial(link = "logit"))

summary(model_3)
summary(model_3rm)
```

In the full dataset, neither temperature nor pressure shows significance toward the probability of O-ring failure, nor does the interaction between these two variables. In the filtered dataset, the situation becomes even more uncertain, as the model couldn't estimate the interaction term, likely due to the small sample size. Overall, incorporating the interaction between temperature and pressure seems to overcomplicate the model, reducing its predictive power compared to the second model. Therefore, the analysis of O-ring failure estimation at 31°F will focus on the second model, where temperature and pressure are considered independently.

## Estimation of O-rings failure with a temperature of 31°F
```{r estimation}
temperature_value <- 31

mean_pressure <- mean(data$Pressure, na.rm = TRUE)
mean_pressurerm <- mean(datarm$Pressure, na.rm = TRUE)

# Since the original report doesn't mention the pressure at take-off, we will use the mean pressure
new_data <- data.frame(Temperature = temperature_value, Pressure = mean_pressure)
new_datarm <- data.frame(Temperature = temperature_value, Pressure = mean_pressurerm)


predicted_prob_m2 <- predict(model_2, newdata = new_data, type = "response")
predicted_prob_m2rm <- predict(model_2rm, newdata = new_datarm, type = "response")


cat("(full dataset) Estimated probability of malfunction at 31°F:", predicted_prob_m2, "\n")
cat("(filtered dataset) Estimated probability of malfunction at 31°F:", predicted_prob_m2rm, "\n")

```

The logistic regression model (temperature + pressure) on the full dataset shows a high probability of O-ring malfunction at 31°F of 68%. Which is consistent with this model showing the significance of the temperature on O-ring failure. However in the filtered dataset, the model showed no significance of the temperature and now shows a probability of failure of 17%. But we have to remember that any successfull launch was removed leading to a model having a strong bias toward failure.


## Final plot
```{r confidence intervals}
# Generate a sequence of temperatures from 30°F to 90°F
tempv <- seq(from = 30, to = 90, by = 0.5)

# Use the mean pressure from the data for prediction
mean_pressure <- mean(data$Pressure, na.rm = TRUE)

# Create a new data frame for predictions with the sequence of temperatures and mean pressure
new_data <- data.frame(Temperature = tempv, Pressure = mean_pressure)

# Predict probabilities using logmodel, model_2, and model_3
predicted_prob_logmodel <- predict(logmodel, newdata = new_data, type = "response")
predicted_prob_model_2 <- predict(model_2, newdata = new_data, type = "response")
predicted_prob_model_3 <- predict(model_3, newdata = new_data, type = "response")

predicted_prob_logmodelrm <- predict(logmodelrm, newdata = new_data, type = "response")
predicted_prob_model_2rm <- predict(model_2rm, newdata = new_data, type = "response")
predicted_prob_model_3rm <- predict(model_3rm, newdata = new_data, type = "response")







# Create a new data frame with predicted probabilities for all models
predicted_data <- data.frame(
  Temperature = rep(tempv, 3),
  PredictedProbability = c(predicted_prob_logmodel, predicted_prob_model_2, predicted_prob_model_3),
  Model = rep(c("logmodel", "model_2", "model_3"), each = length(tempv))
)

predicted_datarm <- data.frame(
  Temperature = rep(tempv, 3),
  PredictedProbability = c(predicted_prob_logmodelrm, predicted_prob_model_2rm, predicted_prob_model_3rm),
  Model = rep(c("logmodel", "model_2", "model_3"), each = length(tempv))
)

# Create the plot
ggplot() +
  # Predicted probabilities as lines for each model
  geom_line(data = predicted_data, aes(x = Temperature, y = PredictedProbability, color = Model), size = 1) +
  # Observed probabilities (Malfunction / Count) as red points
  geom_point(data = data, aes(x = Temperature, y = Malfunction / Count), color = "red", shape = 16) +
  labs(title = "Predicted vs Observed Probability of O-ring Malfunction",
       x = "Temperature (°F)", y = "Probability of Malfunction") +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("blue", "green", "purple")) +
  theme(legend.position = "top")

# Create the plot
ggplot() +
  # Predicted probabilities as lines for each model
  geom_line(data = predicted_datarm, aes(x = Temperature, y = PredictedProbability, color = Model), size = 1) +
  # Observed probabilities (Malfunction / Count) as red points
  geom_point(data = data, aes(x = Temperature, y = Malfunction / Count), color = "red", shape = 16) +
  labs(title = "Predicted vs Observed Probability of O-ring Malfunction",
       x = "Temperature (°F)", y = "Probability of Malfunction") +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("blue", "green", "purple")) +
  theme(legend.position = "top")

```

logmodel = Temperature alone
model_2 = Temperature + Pressure
model_3 = Temperature * pressure
