---
title: "PointingXP"
author: "Fouad Gasmi"
date: "02/01/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r loading libraries}
library(ggplot2)
library(dplyr)

```

## Loading data pre processed by Cornell University Ergonomics website

```{r loading MT data into a dataframe}
meanMTdf <- read.csv("../data/mt_data2.csv", header=T)

```

## Plotting the data together with the linear regression

```{r Plotting the MT data together with the linear regression}
ggplot(meanMTdf, aes(ID, MT)) +
  geom_point() +
  geom_smooth(method='lm')
```
## Linear modelling

```{r linear modeling of MT data}
model <- lm(MT~ID, data = meanMTdf)
summary(model)
```

In this experiment we can see that the ID seems to have an impact on the movement time. However the R²=0.783 find here is different from the R²=0.884 reported by the [experimental software](http://ergo.human.cornell.edu/FittsLaw/FittsLaw.html). The result provided by the website were : MT =132.970 + 180.536log(A/W + 1) with RSquare = 0.884


## Loading data raw data given by Cornell University Ergonomics website
```{r loading raw data into a dataframe}
rawdf <- read.csv("../data/raw_data2.csv", header=T)
```

## Applying Fitt's law formula to get the ID
```{r applying MacKenzie formula}
rawdf$newID <- (log2((rawdf$A / rawdf$W) +1))
```

## Computing the mean MT for each unique combination of A, W, and ID
```{r computing mean time}
newmeanMTdf <- rawdf %>%
  group_by(A, W, newID) %>%
  summarise(mean_MT = mean(MT), .groups = "drop")
```


## Plotting the data together with the linear regression

```{r Plotting the raw data together with the linear regression}
ggplot(newmeanMTdf, aes(newID, mean_MT)) +
  geom_point() +
  geom_smooth(method='lm')
```

## Linear modelling

```{r linear modeling of processed raw data}
model <- lm(mean_MT~newID, data = newmeanMTdf)
summary(model)
```
And now by using the raw data and applying the different operation, we get the same result provided by the webpage
MT = 132.970 + 180.536log(A/W + 1), RSquare = 0.884.


